Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake

### 2014-11-19 paper.tex
Applied Brendan's suggested edits as of Nov 17.

\title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
\author{T. Ben Thompson, Brendan J. Meade -- where to put JS, AP?}
\usepackage{graphicx}
\usepackage[margin=1.0in]{geometry}
\usepackage[nomarkers,figuresonly]{endfloat}
\usepackage{setspace}
\usepackage{natbib}
\doublespacing

@@

\section{Geodetic analysis of Longmen Shan shortening and slip deficit rates}
The geometry of the Beichuan fault and the detachment to the west has been inferred from structural interpretations \citep{Hubbard2010} and seismic reflections observations \citep{Zhang2009, Guo2013}. The Beichuan fault is part of an imbricate thrust system, with fault dips steepening to the northwest. Structural imbrication is typically produced by stacked listric faults that sole to detachments (TODO: CITE?). Furthermore, the large synclinorium parallel to, and west of, the Beichuan fault is consistent with the Beichuan fault shallowing to a detachment based on fault-bend folding theory \citep{Hubbard2010, Suppe1983}. In our model, the range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal ${\sim}20$ km depth where it merges with the detachment surface. At seismogenic depths, a 35$^\circ$ dip has been confirmed via focal mechanisms \citep{Hubbard2010}. Extrapolating from the regional dip of the Sichuan Basin as it extends beneath the Longmenshan range front \citep{Hubbard2010, Guo2013}, the detachment in our model dips ${\sim}1.5^{\circ}$ from an initial depth of 20km to a depth of 23km at a distance of 110km to the northwest. The depth, dip, and extent of the detachment are similar to the best-fitting surface found by \citet{Qi2011} in fitting the co- and post-seismic slip of the Wenchuan earthquake. For our interseismic shortening and slip-deficit models we use an idealized two-dimensional representation of these fault surfaces (Figure \ref{fig:regional_map}).

@@

We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature. The effects may be more pronounced with a shallower fault dip (e.g., many subduction zones). Despite the mediocre effect (${\sim}$15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias. This effect may be even more pronounced for coseismic faulting where deformation is localized in regions of steep topography.

>>>>

\title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
\author{T. Ben Thompson, Brendan J. Meade, Andreas Plesch, John H. Shaw}
\usepackage{graphicx}
\usepackage[margin=1.0in]{geometry}
\usepackage[nomarkers,figuresonly]{endfloat}
\usepackage{setspace}
\usepackage{natbib}
\doublespacing

@@

\section{Geodetic analysis of Longmen Shan shortening and slip deficit rates}
The geometry of the Beichuan fault and the detachment to the west has been inferred from structural interpretations \citep{Hubbard2010} and seismic reflections observations \citep{Zhang2009, Guo2013}. The Beichuan fault is part of an imbricate thrust system, with fault dips steepening to the northwest. Structural imbrication is typically produced by stacked listric faults that sole to detachments. Furthermore, the large synclinorium parallel to, and west of, the Beichuan fault is consistent with the Beichuan fault shallowing to a detachment based on fault-bend folding theory \citep{Hubbard2010, Suppe1983}. In our model, the range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal ${\sim}20$ km depth where it merges with the detachment surface. At 15 km depth, a 35$^{\circ}$ dip is consistent with focal mechanism inversions \citep{Hubbard2010}. Extrapolating from the regional dip of the Sichuan Basin as it extends beneath the Longmenshan range front \citep{Hubbard2010, Guo2013}, the detachment in our model dips ${\sim}1.5^{\circ}$ from an initial depth of 20 km to a depth of 23 km at a distance of 110 km to the northwest. The depth, dip, and extent of the detachment are similar to the best-fitting surface found by \citet{Qi2011} in fitting the co- and post-seismic slip of the Wenchuan earthquake. For our interseismic shortening and slip-deficit models we use an idealized two-dimensional representation of these fault surfaces (Figure \ref{fig:regional_map}).

@@

We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature. Despite the ${\sim}$15\% effect, geodetic analyses in steep regions should consider topographic effects to avoid systematic bias. The effects may be more pronounced with a shallower fault dip or for coseismic displacements where deformation is localized in regions of steep topography.
